\chapter*{Conclusion}
\markleft{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}
\label{chap:conclusion}



During the course of this internship, I addressed several critical challenges and contributed to advancing knowledge in the field of \acrlong{hpo} for \acrlong{llm} \gls{fine_tuning}. My work involved implementing and experimenting with various \acrlong{hpo} algorithms, adapting them to the unique requirements of \gls{fine_tuning}, and conducting comparisons to assess their performance in optimizing expensive black-box functions. Even if results were not groudbreaking, the behavior of the algorithms in front of expensive function, the implementation of \acrfull{hpo} were precious to continuate this work.

Through this, I contributed to bridging the knowledge gap that often limits the accessibility of \gls{fine_tuning} processes for companies and practitioners. Furthermore, I explored the scalability of \acrlong{bo} algorithms in the context of high-performance computing, aligning my work with the long-term objectives of exascale-based projects. This multifaceted approach not only deepened my understanding of the subject but also produced insights and methodologies that can aid future research and practical applications.


Throughout this internship, I gained invaluable knowledge and skills across several domains. I developed a deeper understanding of \acrlong{hpo}, \acrlong{llm} \gls{fine_tuning}, and the intricacies of \acrlong{bo} algorithms, including their theoretical foundations and practical applications. Beyond the technical aspects, I acquired experience in navigating complex research fields, performing thorough literature reviews, and defining well-scoped research problems in collaboration with my supervisor. On a technical level, I honed my programming and development skills, learning to write modular and efficient code while adapting existing frameworks to specific use cases. Moreover, working within resource constraints and leveraging platforms like Grid’5000 strengthened my problem-solving and resource management abilities. 

To conclude this report, I will expand briefly on my professional project, as I introduce since the abstract of this report. One aim of this internship was to decide if I wanted to continue these 6 months with a PhD. And after discussion, a lot of reflection and the opportunity on the right timing, I continue my cursus with a PhD in operation research, at Lille, with a topic on network design for delivery.


